---
title: "Coursera - Reproducible Research - Project #2"
author: "Scott Zwick"
date: "February 14, 2018"
output:
  html_document: default
  pdf_document: default
---
................................................................................................................................................................

##Weather events and their effects on population's health and economic consequenes, based on data from the NOAA Storm Database.

................................................................................................................................................................

###Synopsis
The goal of this report is twofold. First, to present data that shows what types
of storm events are most harmful to public health. Second, to explore which storm
events create the most economic damage. The U.S. National Oceanic and Atmospheric
Administration's (NOAA) storm database provided the data for this study.

```{r setup}
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
knitr::opts_chunk$set(echo = TRUE)
setwd("~/R/Coursera/5-Reproducible Research/Rep_Research_Proj2")
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","repdata-data-StormData.csv.bz2")
strm_data <- read.csv(bzfile("repdata-data-StormData.csv.bz2"), header = TRUE, stringsAsFactors = FALSE)
```

###Data Processing
The documentation for the storm data base indicated there were 48 different storm
events. The database actually included over 900 different storm events. After a
thorough review of the data, a cleansing of the event types was performed to gather
the majority of the events into the specific storm events detailed in the documentation.
When complete, only 11 observations were not classified.

```{r tidy_data1}

## Clean up EVTYPE
strm_data1 <- mutate_all(strm_data, funs(toupper))

strm_data1 <- strm_data %>%
     mutate(EVTYPE = str_to_upper(EVTYPE) %>%
                 str_trim()) %>%
     filter(!str_detect(EVTYPE, "SUMMARY")) %>%
     mutate(EVTYPEFix = str_replace_all(
          EVTYPE,
          c("^(NON|NO)?[ -]?(SEVERE|STRONG|MAJOR|MODERATE|LIGHT|HEAVY|HVY|EXCESSIVE|DENSE|DEEP|DAMAGING|MINOR|EXTREME) " = "",
            "(.*)BLIZZARD(.*)" = "BLIZZARD",
            "HIGH" = "HIGH WIND",
            "FROST" = "FROST/FREEZE",
            "FREEZING FOG STORM" = "FREEZING FOG",
            "WINDCHILL" = "COLD/WIND CHILL",
            "WIND" = "HIGH WIND",
            "COLD" = "COLD/WIND CHILL",
            "TYPHOON" = "HURRICANE (TYPHOON)",
            "FREEZE" = "FROST/FREEZE",
            "ICE" = "ICE STORM",
            "SURF" = "HIGH SURF",
            "AVALANCE" = "AVALANCHE",
            "BLACK ICE" = "FREEZING FOG",
            "BLOWING SNOW(.*)" = "WINTER STORM",
            "(COASTAL|CSTL|BEACH)(.*)FLOOD(.*)" = "COASTAL FLOOD",
            "^HAIL(.*)" = "HAIL",
            "COASTAL ?STORM" = "TROPICAL STORM",
            "BLOW-OUT TIDE(.*)" = "ASTRONOMICAL LOW TIDE",
            "TIDAL(.*)FLOOD(.*)" = "COASTAL FLOOD",
            "TSTM ?WN?D?" = "THUNDERSTORM WIND",
            "TSTM" = "THUNDERSTORM WIND",
            "WND" = "HIGH WIND",
            "TH?UNDEE?R(.*?)WIND(.*)" = "THUNDERSTORM WIND",
            "^THUNDERWIND(.*)" = "THUNDERSTORM WIND",
            "THUNDERWIND(.*)" = "THUNDERSTORM WIND",
            "(.*)(TORNADO|TORNDAO)(.*)" = "TORNADO",
            "TROPICAL STORM(.*)" = "TROPICAL STORM",
            "TORRENTIAL RAIN" = "HEAVY RAIN",
            "UNSEASON([A-Z]*)" = "UNSEASONABLY",
            "&" = "AND",
            "URBAN AND " = "URBAN/",
            "LANDSL(.*)" = "LANDSLIDE",
            "STORM SURGE(.*)" = "STORM SURGE",
            "STREET FLOOD(.*)" = "FLOOD",
            "(.*)URBAN(.*)FL(.*)" = "FLOOD",
            "URBAN/SMALL(.*)" = "FLOOD",
            "(URBAN SMALL)" = "FLOOD",
            "WAY?TER ?SPOUT(.*)" = "WATERSPOUT",
            "WINTER STORM(.*)" = "WINTER STORM",
            "WINTER(.*)MIX" = "WINTER STORM",
            "(.*)WIND CHILL(.*)" = "EXTREME COLD/WIND CHILL",
            "^RIP CURRENT(.*)" = "RIP CURRENT",
            "^RIVER FLOOD(.*)" = "FLOOD",
            "^WIND[S /]{0,}(ADVISORY|DAMAGE|GUSTS|STORM|HAIL)$" = "HIGH WIND",
            "UNUSUALLY|UNUSUAL|UNSEASONABLY" = "UNSEASONABLY",
            "^SNOW(.*)" = "WINTER STORM",
            "WALL CLOUD(.*)" = "FUNNEL CLOUD", 
            "^FLASH FLOOD" = "FLOOD",
            "VOLCAN(.*)" = "VOLCANIC ASH",
            "^ICE JAM(.*)" = "FLOOD",
            "ICE ?/?((AND|SNOW|FLOES|STORM|PELLETS) ?){1,}" = "ICE STORM",
            "LAKE(.*)SNOW" = "LAKE-EFFECT SNOW",
            "LAKE(.*)FLOOD" = "LAKESHORE FLOOD",
            "LATE(.*)SEASON SNOW(.*)" = "HEAVY SNOW",
            "THUNDERSNOW(.*)" = "WINTER SNOW",
            "(THUNDER|THUNDE|THUNER|THUDER)STORM(.*)" = "THUNDERSTORM WIND",
            "SLEET(.*)" = "SLEET",
            "(SMALL STREAM)|(SML STREAM) (FLOODING|FLD|FLOOD)" = "FLOOD",
            "SMALL STREAM(.*)" = "FLOOD",
            "RED FLAG(.*)" = "HIGH SURF",
            "ROUGH (SEAS|SURF)" = "MARINE HIGH WIND",
            "RAIN(.*)" = "HEAVY RAIN",
            "MUD(.*)SLIDE(.*)" = "DEBRIS FLOW",
            "RECORD LOW" = "COLD/WIND CHILL",
            "(NEAR )?RECORD ?" = "",
            "(LIGHTNING|LIGHTING|LIGNTNING)(.*)" = "LIGHTNING",
            "LOCALLY HEAVY " = "",
            "HURRICANE(.*)" = "HURRICANE (TYPHOON)",
            "HIGH WINDS(.*)" = "HIGH WIND",
            "HIGHWAY " = "STREET ",
            "SHOWERS?" = "HEAVY RAIN",
            "WILD(.*)FIRE(.*)" = "WILDFIRE",
            "ROUGH SEAS(.*)" = "MARINE HIGH WIND",
            "ICE(.*)ROAD(.*)" = "ICE STORM",
            "HIGH WIND(.*)" = "HIGH WIND",
            "FREEZING (.*)" = "SLEET",
            "FROST[[:punct:]]{1,}FREEZE" = "FROST/FREEZE",
            "(.*)FUNNEL(.*)" = "FUNNEL CLOUD",
            "FLASH FLOOD(.*)" = "FLASH FLOOD",
            "FLOOD(.*)" = "FLOOD",
            "DRY MIRCOBURST WIND" = "THUNDERSTORM WIND",
            "DRY MICROBURST(.*)" = "THUNDERSTORM WIND",
            "COLD AND(.*)" = "COLD/WIND CHILL",
            "COLD WEATHER" = "COLD/WIND CHILL", 
            "DROUGHT(.*)" = "DROUGHT",
            "/EXCESSIVE HEAT" = "EXCESSIVE HEAT",
            "BRUSH FIRE" = "WILDFIRE",
            "COLD TEMPERATURE" = "COLD/WIND CHILL",
            "COLD WAVE" = "COLD/WIND CHILL",
            "EXTENDED COLD" = "COLD/WIND CHILL",
            "FALLING SNOW/ICE" = "WINTER STORM",
            "DRY MIRCOBURST WIND HIGH WIND" = "THUNDERSTORM WIND",
            "GLAZE/ICE STORM" = "ICESTORM",
            "GLAZE" = "ICE STORM",
            "FOG" = "DENSE FOG",
            "FOG AND COLD/WINDCHILLS" = "COLD/WIND CHILL",
            "GUSTY WIND" = "HIGH WIND",
            "GUSTY WINDS" = "HIGH WIND",
            "HAZARDOUS SURF" = "HIGH SURF",
            "HEAT WAVE" = "HEAT",
            "HEAT WAVE DROUGHT" = "HEAT",
            "HEAT WAVES" = "HEAT",
            "HEAVY HEAVY RAIN" = "HEAVY RAIN",
            "COLD/WINDS" = "COLD/WIND CHILL",
            "DENSE FOG AND COLD/WIND CHILLS" = "DENSE FOG",
            "FREZZING DENSE FOG" = "FREEZING FOG",
            "HEAT DROUGHT" = "HEAT",
            "HEATS" = "HEAT",
            "HIGH SWELLS" = "MARINE HIGH WIND",
            "HIGH SEAS" = "MARINE HIGH WIND",
            "HIGH WATER" = "FLOOD",
            "HIGH WAVES" = "MARINE HIGH WIND",
            "HIGH SURF" = "MARINE HIGH WIND",
            "HIGH WINDS" = "HIGH WIND",
            "HYPERTHERMIA/EXPOSURE" = "EXTREME COLD/WIND CHILL",
            "HYPOTHERMIA/EXPOSURE" = "EXTREME COLD/WIND CHILL",
            "HYPOTHERMIA" = "EXTREME COLD/WIND CHILL",
            "ICE STORM/FLASH FLOOD" = "ICE STORM",
            "ICESTORM" = "ICE STORM",
            "ICY ROADS" = "ICE STORM",
            "LANDSLIDE" = "DEBRIS FLOW",
            "LOW TEMPERATURE" = "COLD/WIND CHILL",
            "MARINE STRONG WIND" = "MARINE HIGH WIND",
            "MARINE WIND" = "MARINE STRONG WIND",
            "RAPIDLY RISING WATER" = "FLASH FLOOD",
            "SMALL HAIL" = "HAIL",
            "STORM SURGE" = "STORM SURGE/TIDE",
            "SURF AND WIND" = "HIGH SURF",
            "SURF/MARINE HIGH WIND" = "MARINE HIGH WIND",
            "UNSEASONABLY COLD" = "COLD/WIND CHILL",
            "UNSEASONABLY WARM" = "HEAT",
            "UNSEASONABLY WARM AND DRY" = "DROUGHT",
            "WARM WEATHER" = "HEAT",
            "HEAT AND DRY" = "DROUGHT",
            "NON THUNDERSTORM WIND" = "HIGH WIND",
            "ROGUE WAVE" = "MARINE HIGH WIND",
            "WHIRLWIND" = "DUST DEVIL",
            "WINDS" = "HIGH WIND",
            "WINTER SNOW" = "WINTER STORM",
            "WINTRY MIX" = "WINTER WEATHER",
            "MARINE ACCIDENT" = "MARINE THUNDERSTORM WIND",
            "MARINE MISHAP" = "MARINE THUNDERSTORM WIND",
            "HAZARDOUS MARINE HIGH WIND" = "MARINE HIGH WIND",
            "HIGH MARINE HIGH WIND" = "MARINE HIGH WIND",
            "FROST/FROST/FREEZE" = "FROST/FREEZE",
            "MARINE HIGH WIND AND WIND" = "MARINE HIGH WIND",
            "MARINE HIGH WIND AND HIGH WIND" = "MARINE HIGH WIND",
            "MARINE HIGH WIND/MARINE HIGH WIND" = "MARINE HIGH WIND",
            "MARINE STRONG HIGH WIND" = "MARINE HIGH WIND",
            "ROUGH MARINE HIGH WIND" = "MARINE HIGH WIND",
            "SEAS" = "MARINE HIGH WIND",
            "WINTER STORM STORM" = "WINTER STORM",
            "GUSTY HIGH WIND" = "HIGH WIND",
            "HIGH HIGH WIND" = "HIGH WIND",
            "DRY MIRCOBURST HIGH WIND" = "THUNDERSTORM WIND",
            "WHIRLHIGH WIND" = "HIGH WIND",
            "OTHER" = "DUST DEVIL",
            "MARINE HIGH WIND/HIGH WIND" = "MARINE HIGH WIND",
            "MARINE TSTM WIND" = "MARINE THUNDERSTORM WIND",
            "TSTM ?WN?D?" = "THUNDERSTORM WIND",
            "TSTM" = "THUNDERSTORM",
            "COLD/(.*)" = "COLD",
            "AGRICULTURAL FREEZE" = "FROST/FREEZE",
            "APACHE COUNTY" = "HIGH WIND",
            "ASTRONOMICAL HIGH TIDE" = "HIGH SURF",
            "BEACH EROSION" = "HIGH SURF",
            "BLOWING DUST" = "DUST STORM",
            "BLOWING SNO(.*)" = "WINTER STORM",
            "BREAKUP FLOOD" = "FLOOD",
            "COASTAL EROSION" = "HIGH WIND",
            "COASTAL FLOOD" = "FLOOD",
            "COASTAL SURGE" = "STORM SURGE/TIDE",
            "COASTAL STORM" = "MARINE THUNDERSTORM WIND",
            "COOL AND WET" = "COLD/WIND CHILL",
            "DAM(.*)" = "FLASH FLOOD",
            "DOWNBURST" = "HEAVY RAIN",
            "DUST DEVIL WATERSPOUT" = "DUST DEVIL",
            "DUST STORM/WIND" = "DUST STORM",
            "EROSION/FLOOD" = "FLOOD",
            "EXTENDED COLD" = "COLD/WIND CHILL",
            "FOREST FIRES" = "WILDFIRE",
            "FREEZING PRECIP" = "SLEET",
            "GLAZE(.*)" = "SLEET",
            "EARLY FROST" = "FROST/FREEZE", 
            "EXTREME COLD" = "EXTREME COLD/WIND CHILL",
            "EXTREME HEAT" = "EXCESSIVE HEAT",
            "EXTREME WINDCHILL" = "EXTREME COLD/WIND CHILL",
            "GRADIENT WIND" = "HIGH WIND",
            "GRASS FIRES" = "WILDFIRE",
            "GUSTNADO" = "THUNDERSTORM WIND",
            "HARD FREEZE" = "FROST/FREEZE",
            "HEAVY LAKE EFFECT SNOW" = "LAKE-EFFECT SNOW",
            "HEAVY MIX" = "WINTER STORM",
            "HEAVY PRECIPITATION" = "HEAVY RAIN",
            "HEAVY SNOW AND STRONG WINDS" = "BLIZZARD",
            "HEAVY SNOW (.*)" = "WINTER STORM",
            "HEAVY SNOW/(.*)" = "HEAVY SNOW",
            "HEAVY SNOWPACK" = "HEAVY SNOW",
            "EXCESSIVE SNOW" = "HEAVY SNOW",
            "HEAVY SNOW-SQUALLS" = "HEAVY SNOW",
            "HEAVY SURF(.*)" = "HIGH SURF",
            "HEAVY SWELLS" = "HIGH SURF",
            "HIGH SURF ADVISORY" = "HIGH SURF",
            "HIGH TIDES" = "STORM SURGE/TIDE",
            "ICE/STRONG WINDS" = "ICE STORM",
            "LAKE EFFECT SNOW" = "LAKE-EFFECT SNOW",
            "LAKE FLOOD" = "FLOOD",
            "LANDSPOUT" = "DUST DEVIL",
            "MICROBURST(.*)" = "THUNDERSTORM WIND",
            "MIXED PRECIPITATION" = "WINTER WEATHER",
            "MUDSLIDE" = "DEBRIS FLOW",
            "NON-THUNDERSTORM" = "HIGH WIND",
            "RIVER AND STREAM FLOOD" = "FLOOD",
            "ROCK SLIDE" = "DEBRIS FLOW", 
            "RURAL FLOOD" = "FLOOD",
            "SMALL STREAM FLOOD" = "FLOOD",
            "STORM FORCE WINDS" = "HIGH WIND",
            "THUNDERSNOW" = "WINTER STORM",
            "TIDAL FLOOD" = "STORM SURGE/TIDE",
            "URBAN FLOOD" = "FLOOD",
            "WET THUNDERSTORM WIND" = "THUNDERSTORM WIND",
            "WIND AND WAVE" = "HIGH WIND",
            "UNSEASONABLY RAIN" = "HEAVY RAIN",
            "^COLD$" = "COLD/WIND CHILL",
            "^FOG$" = "DENSE FOG", 
            "^FREEZE$" = "FROST/FREEZE",
            "^FROST$" = "FROST/FREEZE",
            "^ICE$" = "ICE STORM",
            "^MARINE THUNDERSTORM$" = "MARINE THUNDERSTORM WIND",
            "^RAIN$" = "HEAVY RAIN",
            "^SNOW$" = "HEAVY SNOW",
            "^STORM SURGE$" = "STORM SURGE/TIDE",
            "^THUNDERSTORM$" = "THUNDERSTORM WIND",
            "^TYPHOON$" = "HURRICANE (TYPHOON)",
            "^WIND$" = "HIGH WIND",
            "^WIND CHILL$" = "COLD/WIND CHILL",
            "^WINDS$" = "HIGH WIND",
            "ABNORMAL WARMTH" = "HEAT",
            "ABNORMALLY DRY" = "DROUGHT",
            "ABNORMALLY WET" = "HEAVY RAIN",
            "ACCUMULATED SNOWFALL" = "HEAVY SNOW",
            "AGRICULTURAL FROST/FREEZE" = "FROST/FREEZE",
            "ASTRONOMICAL HIGH WIND" = "HIGH WIND",
            "BEACH EROSIN" = "HIGH SURF",
            "BELOW NORMAL PRECIPITATION" = "DROUGHT",
            "COOL(.*)" = "COLD/WIND CHILL",
            "DRIEST MONTH" = "DROUGHT",
            "DRIFTING SNOW" = "BLIZZARD",
            "DRY(.*)" = "DROUGHT",
            "DUST DEVEL" = "DUST DEVIL",
            "DUST STORM/HIGH WIND" = "DUST STORM",
            "DUSTSTORM" = "DUST STORM",
            "EARLY HEAVY RAIN" = "HEAVY RAIN",
            "EARLY SNOW(.*)" = "HEAVY SNOW",
            "EXCESSIVELY DROUGHT" = "DROUGHT",
            "EXTREMELY WET" = "HEAVY RAIN",
            "FIRST FROST/FREEZE" = "FROST/FREEZE",
            "FIRST SNOW" = "HEAVY SNOW",
            "FLASH FLOOODING" = "FLASH FLOOD",
            "FROST/FREEZE(.*)" = "FROST/FREEZE",
            "GRADIENT HIGH WIND" = "HIGH WIND",
            "GUSTY LAKE HIGH WIND" = "HIGH WIND",
            "GUSTY THUNDERSTORM WIND" = "THUNDERSTORM WIND",
            "HARD FROST/FREEZE" = "FROST/FREEZE",
            "HEAT(.*)" = "HEAT",
            "HEAVY RAIN HIGH WIND" = "HEAVY RAIN",
            "HIGH WIND WIND" = "HIGH WIND",
            "HOT(.*)" = "HEAT",
            "ICE STORM(.*)" = "ICE STORM",
            "LATE FROST/FREEZE" = "FROST/FREEZE",
            "LATE MARINE HIGH WINDON HAIL" = "MARINE HIGH WIND",
            "LATE SNOW" = "HEAVY SNOW",
            "LOCAL FLASH FLOOD" = "FLASH FLOOD",
            "LOCAL FLOOD" = "FLOOD",
            "LOW HEAVY RAIN" = "HEAVY RAIN",
            "MARINE HIGH WIND(.*)" = "MARINE HIGH WIND",
            "MAY SNOW" = "HEAVY SNOW",
            "MILD AND DROUGHT" = "DROUGHT",
            "MILD/DROUGHT" = "DROUGHT",
            "MIX" = "WINTER WEATHER",
            "MONTHLY HEAVY RAIN" = "HEAVY RAIN",
            "MONTHLY PRECIPITATION" = "HEAVY RAIN",
            "MONTHLY SNOWFALL" = "HEAVY SNOW",
            "MOUNTAIN SNOWS" = "HEAVY SNOW",
            "NORMAL PRECIPITATION" = "HEAVY RAIN",
            "PATCHY DENSE DENSE FOG" = "DENSE FOG",
            "PATCHY ICE STORM" = "ICE STORM",
            "PROLONG WARMTH" = "HEAT",
            "PROLONGED HEAVY RAIN" = "HEAVY RAIN",
            "REMNANTS OF FLOYD" = "HURRICANE (TYPHOON)",
            "SAHARAN DUST" = "DUST STORM",
            "^SMOKE$" = "DENSE SMOKE",
            "SNOWFALL" = "HEAVY SNOW",
            "STORM FORCE HIGH WIND" = "HIGH WIND",
            "STREAM FLOOD" = "FLOOD",
            "THUNDERSTORM WIND AND" = "THUNDERSTORM WIND",
            "UNMARINE HIGH WIN(.*)" = "MARINE HIGH WIND",
            "VERY DROUGHT" = "DROUGHT",
            "VERY WARM" = "HEAT",
            "VOG" = "DENSE FOG",
            "WARM(.*)" = "HEAT",
            "WET SNOW" = "HEAVY SNOW",
            "WET(.*)" = "HEAVY RAIN",
            "WILDFIRES" = "WILDFIRE",
            "MONTHLY TEMPERATURE" = "COLD/WIND CHILL"
          )
     ))
strm_data1$EVTYPEFix <- gsub("WINTER WEATHERED PRECIP", "WINTER WEATHER", strm_data1$EVTYPEFix)
strm_data1$EVTYPEFix <- gsub("HEAVY HEAVY SNOW", "HEAVY SNOW", strm_data1$EVTYPEFix)
```

### 1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

```{r fat+inj}
## Create column with the sum of fatalities and injuries
strm_data1$FAT_INJ <- strm_data1$FATALITIES + strm_data1$INJURIES

## Get the sum of the fatalities/injuries by EVTYPE
fat_inj <- select(strm_data1, EVTYPEFix, FAT_INJ)
fat_inj_tot <- aggregate(FAT_INJ~EVTYPEFix, fat_inj, sum)

## Get top 5
fat_inj_5 <- fat_inj_tot[order(fat_inj_tot$FAT_INJ, decreasing=T)[1:5],]

## Bar graph
barplot(fat_inj_5$FAT_INJ, main = "Total Fatalities and Injuries", 
        names.arg = fat_inj_5$EVTYPEFix, cex.names = .75)
```

Per the data collected, tornados have caused the most direct injuries and fatalities during the study period. 


## 2. Across the United States, which types of events have the greatest economic consequences?
###Data Processing
PROPDMGEXP and CROPDMGEXP, used as multipliers for property and crop damabe have
several non-standard values not defined in the storm database documentation. 
Elimination of the non-standard EXP types resulted in eliminating 70 of the 
902,151 rows, or .01% of the data.
```{r calc_max}

strm_data2 <- strm_data1[!(strm_data1$PROPDMGEXP == "2" |
                                strm_data1$PROPDMGEXP == "3" |
                                strm_data1$PROPDMGEXP == "4" |
                                strm_data1$PROPDMGEXP == "5" |
                                strm_data1$PROPDMGEXP == "6" |
                                strm_data1$PROPDMGEXP == "7" |
                                strm_data1$PROPDMGEXP == "-" |
                                strm_data1$PROPDMGEXP == "+" |
                                strm_data1$PROPDMGEXP == "H"
),]

strm_data2 <- strm_data2[!(strm_data2$CROPDMGEXP == "?"),]

## Adjusted EXP to a multiplier
strm_data2$PROPDMGEXP <- gsub("K", "1000", strm_data2$PROPDMGEXP)
strm_data2$PROPDMGEXP <- gsub("M", "1000000", strm_data2$PROPDMGEXP)
strm_data2$PROPDMGEXP <- gsub("B", "1000000000", strm_data2$PROPDMGEXP)

strm_data2$CROPDMGEXP <- gsub("K", "1000", strm_data2$CROPDMGEXP)
strm_data2$CROPDMGEXP <- gsub("M", "1000000", strm_data2$CROPDMGEXP)
strm_data2$CROPDMGEXP <- gsub("B", "1000000000", strm_data2$CROPDMGEXP)

## Convert column to numeric
conv_cols <- c("PROPDMG","PROPDMGEXP", "CROPDMG", "CROPDMGEXP")
strm_data2[conv_cols] <- sapply(strm_data2[conv_cols],as.numeric)

## Create damage totals by muliplying DMG columns by EXP columns
strm_data2$PROPDMG_tot <- strm_data2$PROPDMG * strm_data2$PROPDMGEXP
strm_data2$CROPDMG_tot <- strm_data2$CROPDMG * strm_data2$CROPDMGEXP

## Add Property Damage to Crop Damage
strm_data2$DMG_tot <- strm_data2$PROPDMG_tot + strm_data2$CROPDMG_tot
DMG_total <- select(strm_data2, EVTYPEFix, DMG_tot)

## Get the sum of the damages by EVTYPE
DAMAGE_total <- aggregate(DMG_tot~EVTYPEFix, DMG_total, sum)

## Get top 5
propcrop_5 <- DAMAGE_total[order(DAMAGE_total$DMG_tot, decreasing=T)[1:5],]

## Bar graph
barplot(propcrop_5$DMG_tot, main = "Total Dollars of Property and Crop Damages", 
        names.arg = propcrop_5$EVTYPEFix, cex.names = .65)
```

Per the database, floods have caused the highest amount of property and crop damage during the study period. 